HO unification from object language to meta language

### 2024-04-03 paper/main.tex
tex: beta update

In order to
encompass\todo[inline]{not sure has the right meaning}
this unification, term compilation should capture the
terms $t$ in \maybebeta and replace them with fresh varaibles $v$.
\todo[inline]{we use capitals for unif variables}
As per
\maybebeta, the variables $v$ and the terms $t$ are linked through a \linkbeta.

@@

\begin{invariant}[\linkbeta \rhs]
  The \rhs of any \linkbeta has the shape $X~t_1\dots t_n$ such that $X$ is a
  flexible term and $t_1\dots t_n$ is not in \llambda.
\end{invariant}

\linkbeta are put in \linkStore and activated when \rhs falls in \llambda. 
\todo[inline]{un po' ridondante con 8.2, non so se serva}

@@

\begin{proof}[Proof sketch]
  Given $X \appsep t_1\dots t_n \Ue t$ where $t$ is a rigid term and
  $t_1\dots t_n$ is not in \llambda. By construction, $X \appsep t_1\dots t_n$
  is replaced with a variable $V$, and the \linkbeta \linkbetaM{\Gamma}{V}{X
  \appsep t_1\dots t_n} is created. The unification instantiates $V$ to $t$,
  making the \lhs of the link a rigid term, while \rhs is still in \maybebeta.
  The original problem is in fact outside \llambda and unification fails.
\end{proof}

@@

Detection of \maybebeta is quite simple to implement in the compiler, since it
is sufficient to capture applications with flexible head and argument that
are not in \llambda.

@@

In the following, we pose $X_{s_1\dots s_n}$ to represent a variable $X$ with
the list of distinct names $s_1\dots s_n$ as scope. Moreover, $X_{s_1\dots
s_n}\appsep t_1\dots t_m$ is the application of that variable to the list of
terms $t_1\dots t_m$. Note that \rhsBeta is equivalent to \elpiIn{app[uva N S |
T]}.

>>>>

To address
this unification problem, term compilation should capture the
terms $t$ in \maybebeta and replace them with fresh varaibles $X$.
The variables $X$ and the terms $t$ are linked through a \linkbeta.

@@

\begin{invariant}[\linkbeta \rhs]
  The \rhs of any \linkbeta has the shape \rhsBeta such that $X$ is a
  unification variable with scope $s_1\dots s_n$ and $t_1\dots t_m$ is a list of
  terms. This is equivalent to \elpiIn{app[uva X S | L]}, where \elpiIn{S}~$=
  s_1\dots s_n$ and \elpiIn{L}~$=t_1\dots t_m$.
\end{invariant}

% \linkbeta are put in \linkStore and activated when \rhs falls in \llambda. 
% \todo[inline]{un po' ridondante con 8.2, non so se serva}

@@

\begin{proof}[Proof sketch]
  Given $X \appsep t_1\dots t_n \Ue t$ where $t$ is a rigid term and
  $t_1\dots t_n$ is not in \llambda. By construction, $X \appsep t_1\dots t_n$
  is replaced with a variable $Y$, and the \linkbeta \linkbetaM{\Gamma}{Y}{X
  \appsep t_1\dots t_n} is created. The unification instantiates $Y$ to $t$,
  making the \lhs of the link a rigid term, while \rhs is still in \maybebeta.
  The original problem is in fact outside \llambda.
\end{proof}

@@

Detection of \maybebeta is quite simple to implement in the compiler, since it
is sufficient to detect applications with flexible head and argument that
are not in \llambda.

@@

% In the following, we pose $X_{s_1\dots s_n}$ to represent a variable $X$ with
% the list of distinct names $s_1\dots s_n$ as scope. Moreover, $X_{s_1\dots
% s_n}\appsep t_1\dots t_m$ is the application of that variable to the list of
% terms $t_1\dots t_m$. Note that \rhsBeta is equivalent to \elpiIn{app[uva N S |
% T]}.
